---
title: "Getting Started with Leaflet"
author: "Frida Lützow Kirk, Ida-Bella Bach Greve & Kristine Xenia Storgaard Søndberg"
date: "`r Sys.Date()`"
output: html_document
---

## Install Required Packages

```{r install-packages, eval=FALSE}
install.packages("leaflet")
install.packages("htmltools")
install.packages("htmlwidgets")
install.packages("tidyverse")
install.packages("googlesheets4")
```

## Load Libraries

```{r setup}
library(leaflet)
library(htmltools)
library(htmlwidgets)
library(tidyverse)
library(googlesheets4)
```

---

# Example 1: Create a Leaflet Map of Europe

Learn to create a map by plotting three point markers called Robin, Jakub, and Jannes.

```{r example1}
# Create labels for your points
labels <- c("Robin", "Jakub", "Jannes")

# Create the map
leaflet() %>%
  addProviderTiles("Esri.WorldPhysical") %>%
  addAwesomeMarkers(
    lng = c(-3, 23, 11),
    lat = c(52, 53, 49),
    popup = labels
  )
```

---

# Example 2: Create a Leaflet Map of Sydney

The `setView()` function helps focus and zoom the map to a specific area.

```{r example2}
leaflet() %>%
  addTiles() %>%
  addProviderTiles(
    "Esri.WorldImagery",
    options = providerTileOptions(opacity = 0.5)
  ) %>%
  setView(lng = 151.105006, lat = -33.8767231, zoom = 12)
```

## Question 1

What is the order of longitude and latitude in the `setView()` function?

### Answer

1. longtitude, 2. latitude

---

# Multiple Background Layers

```{r europe-layers}
leaflet() %>%
  addTiles() %>%
  setView(lng = 2.34, lat = 48.85, zoom = 5) %>%
  addProviderTiles("Esri.WorldPhysical", group = "Physical") %>%
  addProviderTiles("Esri.WorldImagery", group = "Aerial") %>%
  addProviderTiles("MtbMap", group = "Geo") %>%
  addLayersControl(
    baseGroups = c("Geo", "Aerial", "Physical"),
    options = layersControlOptions(collapsed = F)
  )
```

## Question 2

How does the map above change if you replace the T in the last line of code above with F?

### Answer

The layer control menu changes from being collapsed (hidden) to being expanded (visible) by default.

---

# Example 3: Sydney Harbour Display with Multiple Layers

## Step 1: Create a Base Map

```{r sydney-base}
l_aus <- leaflet() %>%
  setView(151.2339084, -33.85089, zoom = 13)

l_aus
```

## Step 2: Load Esri Background Layers

```{r esri-layers}
esri <- grep("^Esri", providers, value = TRUE)
esri
```

```{r add-esri-layers}
for (provider in esri) {
  l_aus <- l_aus %>%
    addProviderTiles(provider, group = provider)
}

l_aus
```

## Step 3: Build the Final Sydney Map

```{r aus-map}
AUSmap <- l_aus %>%
  addLayersControl(
    baseGroups = names(esri),
    options = layersControlOptions(collapsed = FALSE)
  ) %>%
  addMiniMap(
    tiles = esri[[1]],
    toggleDisplay = TRUE,
    position = "bottomright"
  ) %>%
  addMeasure(
    position = "bottomleft",
    primaryLengthUnit = "meters",
    primaryAreaUnit = "sqmeters",
    activeColor = "#3D535D",
    completedColor = "#7D4479"
  ) %>%
  htmlwidgets::onRender(
    "
    function(el, x) {
      var myMap = this;
      myMap.on('baselayerchange', function (e) {
        myMap.minimap.changeLayer(L.tileLayer.provider(e.name));
      })
    }
    "
  ) %>%
  addControl("", position = "topright")

AUSmap
```

## Save the Map

```{r save-aus-map, eval=FALSE}
saveWidget(AUSmap, "AUSmap.html", selfcontained = TRUE)
```

---

# Task 1: Create a Danish Equivalent of AUSmap

```{r danmap}
l_dk <- leaflet() %>%
  setView(10.204221884316414, 56.16067847421799, zoom = 15)

for (provider in esri) {
  l_dk <- l_dk %>%
    addProviderTiles(provider, group = provider)
}

DANmap <- l_dk %>%
  addLayersControl(
    baseGroups = names(esri),
    options = layersControlOptions(collapsed = FALSE)
  ) %>%
  addMiniMap(
    tiles = esri[[1]],
    toggleDisplay = TRUE,
    position = "bottomright"
  ) %>%
  addMeasure(
    position = "bottomleft",
    primaryLengthUnit = "meters",
    primaryAreaUnit = "sqmeters",
    activeColor = "#3D535D",
    completedColor = "#7D4479"
  ) %>%
  htmlwidgets::onRender(
    "
    function(el, x) {
      var myMap = this;
      myMap.on('baselayerchange', function (e) {
        myMap.minimap.changeLayer(L.tileLayer.provider(e.name));
      })
    }
    "
  ) %>%
  addControl("", position = "topright")

DANmap
```

## Save DANmap

```{r save-danmap, eval=FALSE}
saveWidget(DANmap, "DANmap.html", selfcontained = TRUE)
```

---

# Add Data to Leaflet

## Create Coordinate Data

Go to:

- https://bit.ly/CreateCoordinates1

Enter coordinates of your favorite leisure places in Denmark.

## Read the Google Sheet into R

```{r deauth}
gs4_deauth()
```

```{r read-sheet}
places <- read_sheet(
  "https://docs.google.com/spreadsheets/d/1PlxsPElZML8LZKyXbqdAYeQCDIvDps2McZx1cTVWSzI/edit#gid=124710918",
  col_types = "cccnncnc",
  range = "DAM2026"
)

places
```

## Check the Data

```{r check-data}
glimpse(places)

places %>%
  filter(!is.na(Longitude)) %>%
  filter(!is.na(Latitude))
```

## Question 3

Are the Latitude and Longitude columns present? Do they contain numeric decimal degrees?

### Answer

Yes, there are numeric decimal degree coordinates because the filter function confirms valid longitude and latitude values.

---

# Create a Basic Student Map

```{r student-map}
studentmap <- leaflet() %>%
  addTiles() %>%
  addMarkers(
    lng = places$Longitude,
    lat = places$Latitude,
    popup = paste(places$Description, "<br>", places$Type)
  )

studentmap
```

## Save Student Map

```{r save-studentmap, eval=FALSE}
saveWidget(studentmap, "studentmap.html", selfcontained = TRUE)
```

---

# Task 2: Add Google Sheet Data to DANmap

```{r task2}
task2map <- DANmap %>%
  addMarkers(
    lng = places$Longitude,
    lat = places$Latitude,
    popup = paste(places$Description, "<br>", places$Type)
  )

task2map
```

---

# Task 3: Cluster the Points in Leaflet

```{r task3}
task3map <- DANmap %>%
  addMarkers(
    lng = places$Longitude,
    lat = places$Latitude,
    popup = paste(places$Description, "<br>", places$Type),
    clusterOptions = markerClusterOptions()
  )

task3map
```

---

# Task 4: Compare Clustered vs Non-Clustered Maps

## Answer

The clustering effect creates a cleaner overview of where the hotspots are located, while the individual markers provide more precise locations immediately.

---

# Task 5: Display Notes and Classifications in the Map

You can include additional columns in the popup text.

```{r task5}
DANmap %>%
  addMarkers(
    lng = places$Longitude,
    lat = places$Latitude,
    popup = paste(
      places$Description,
      "<br>",
      places$Type,
      "<br>",
      places$Notes
    )
  )
```

---

# Congratulations

You have completed the Leaflet workshop.

